Supplementary Materials for \\ Modeling the competing effects of the immune system and EMT on epithelial cancers

### 2019-09-16 Paper/Supplement/Supplement_rewrite.tex
Rewrote after referees got back to us

\title{Supplementary Materials for \\
Modeling the competing effects of the immune system and EMT on tumor development
}
\author{Daniel R. Bergman$^{1}$,
Matthew K. Karikomi\,$^{1}$,
Qing Nie\,$^{1,2,*}$
and \\Adam L. MacLean\,$^{3,*}$
}
\affil{
  $^1$Department of Mathematics, University of California, Irvine,  Irvine, CA 92697, USA \\
  $^2$Department of Cell and Developmental Biology, University of California, Irvine, Irvine, CA 92697, USA \\
  $^3$Department of Biological Sciences, University of Southern California, Los Angeles, CA 90089, USA \\
  $^*$Correspondence:  qnie@uci.edu (Q.N.); macleana@usc.edu (A.L.M.)
}
\renewcommand\Authands{ and }

@@

\begin{table}[H]
\begin{center}
 \begin{tabular}{|| c | c||} 
 \hline
 Name & Description  \\ [0.5ex] 
 \hline
 $p$ & proliferation rate of tissue cells \\ 
 \hline
 $d_C$  & death rate of tissue cells \\
 \hline
$\Delta_\text{MIE}$ &  mesenchymal immune evasion \\
 \hline
 $\Delta_\text{MGA}$ & mesenchymal growth arrest    \\
 \hline
  $\Delta_A$ & mutant cells decreased apoptosis  \\
  \hline
  $\Delta_\text{IE}$ & mutant cells increased immune evasion  \\
  \hline
  $\Delta_P$ & mutant cells increased proliferation  \\
  \hline
 $K_0$ & EC50 term for negative feedback of tissue cells on own proliferation\\
 \hline
 $K_1$ & EC50 term for probability of NK cell finding mutant cell\\
 \hline
  $K_2$ & EC50 term for Treg inhibition of cytotoxic functions  \\
  \hline
  $K_3$ & EC50 term for how much TGF-$\beta$ each cell has \\
  \hline
 $E_\text{NK}$ & rate of NKs clearing mutants  \\
  \hline
  $E_\text{CTL}$ & rate of CTLs clearing mutants \\
  \hline
  $\sigma_\text{NK}$ & NK source rate \\ 
  \hline
  $\sigma_\text{CTL}$ & CTL source rate per cleared mutant cell \\ 
  \hline
  $\sigma_\text{Treg}$ & Treg source rate per cleared mutant cell \\ 
  \hline
  $k_\text{EMT}$ & EMT/MET rate  \\
  \hline
  $\sigma$ & standard deviation of noise in TGF-$\beta$ each cell receives  \\
  \hline
 $\tau_\text{max}$ & max amount of TGF-$\beta$ any cell can receive \\
  \hline 
 $\tau_\text{MUT}$ & rate of TGF-$\beta$ production by mutant cells\\
  \hline
 $\tau_\text{Treg}$ & rate of TGF-$\beta$ production by Treg\\
  \hline
\end{tabular}
  \caption{The model parameter names and descriptions. Note that many of these values are affected by the inflammation state of the system.}
\end{center}
\end{table}

@@

\begin{table}[H]
\begin{center}
 \begin{tabular}{|| c | p{7cm} | c | c ||} 
 \hline
 Name & Description & INFL Low Value & INFL High Value  \\ [0.5ex] 
 \hline\hline
  $p$ & proliferation rate of tissue cells & 0.28 & \\ 
 \hline
 $d_C$  & death rate of tissue cells & 0.14 & \\
 \hline
$\Delta_\text{MIE}$ &  MIE & 0.6 & \\
 \hline
 $\Delta_\text{MGA}$ & MGA  & 0.2 &  \\
 \hline
  $\Delta_A$ & mutant cells decreased apoptosis & 0.3 &  \\
  \hline
  $\Delta_\text{IE}$ & mutant cells increased immune evasion & 0.48 &  \\
  \hline
  $\Delta_P$ & mutant cells increased proliferation & 0.36 &  \\
  \hline
 $K_0$ & EC50 term for negative feedback of tissue cells on own proliferation & 80 & \\
 \hline
 $K_1$ & EC50 term for probability of NK cell finding mutant cell & 8 & \\
 \hline
  $K_2$ & EC50 term for Treg inhibition of cytotoxic functions & 5 & 0.025  \\
  \hline
  $K_3$ & EC50 term for how much TGF-$\beta$ each cell has & 200 & \\
  \hline
 $E_\text{NK}$ & rate of NKs clearing mutants & 10 & 30  \\
  \hline
  $E_\text{CTL}$ & rate of CTLs clearing mutants & 200 & 600 \\
  \hline
  $\sigma_\text{NK}$ & NK source rate & 1.3 &  \\ 
  \hline
  $\sigma_\text{CTL}$ & CTL source rate per cleared mutant cell & 100 & \\ 
  \hline
  $\sigma_\text{Treg}$ & Treg source rate per cleared mutant cell & 200 & \\ 
  \hline
  $k_\text{EMT}$ & EMT/MET rate & 0.01 & \\
  \hline
  $\sigma$ & standard deviation of noise in TGF-$\beta$ each cell receives & 6 & \\
  \hline
 $\tau_\text{max}$ & max amount of TGF-$\beta$ any cell can receive & 500 &\\
  \hline 
 $\tau_\text{MUT}$ & rate of TGF-$\beta$ production by mutant cells & 0.05 & \\
  \hline
 $\tau_\text{Treg}$ & rate of TGF-$\beta$ production by Treg & 0.5 & \\
  \hline
  & RP Cancer Line  & 0.5 &  \\ 
 \hline
&  INFL High Duration & 30&   \\
 \hline
 & INFL Low Duration &30 &   \\
 \hline
& Mes Threshold & 0.7 &    \\
 \hline
  & maximum initial mutation damage after warmup  & 0.01 &\\
  \hline
  & increase in probability to mutate for non-mutating proliferating cells & 0.0001 & \\
  \hline
\end{tabular}
  \caption{The model parameter names, descriptions, and values during both low and high inflammation. Parameters with only one value do not change with the inflammatory state.}
\end{center}
\end{table}

>>>>

\title{Supplementary Materials for \\
Modeling the competing effects of the immune system and EMT on epithelial cancers
}
\author{Daniel R. Bergman$^{1}$,
Matthew Karikomi\,$^{1}$,
Qing Nie\,$^{1,2,*}$
and \\Adam L. MacLean\,$^{3,*}$
}
\affil{
  $^1$Department of Mathematics, University of California, Irvine,  Irvine, CA 92697, USA \\
  $^2$Department of Cell and Developmental Biology, University of California, Irvine, Irvine, CA 92697, USA \\
  $^3$Department of Biological Sciences, University of Southern California, Los Angeles, CA 90089, USA \\
  $^*$Correspondence:  qnie@uci.edu (Q.N.); macleana@usc.edu (A.L.M.)
}
\renewcommand\Authands{ and }

@@

\begin{table}[H]
\begin{center}
 \begin{tabular}{|| c | c||} 
 \hline
 Name & Description  \\ [0.5ex] 
 \hline
 $p$ & proliferation rate of tissue cells \\ 
 \hline
 $d_C$  & death rate of tissue cells \\
 \hline
$\Delta_\text{MIE}$ &  mesenchymal immune evasion \\
 \hline
 $\Delta_\text{MGA}$ & mesenchymal growth arrest    \\
 \hline
  $\Delta_A$ & mutant cells decreased apoptosis  \\
  \hline
  $\Delta_\text{IE}$ & mutant cells increased immune evasion  \\
  \hline
  $\Delta_P$ & mutant cells increased proliferation  \\
  \hline
 $K_0$ & EC50 term for negative feedback of tissue cells on own proliferation\\
 \hline
 $K_1$ & EC50 term for probability of NK cell finding mutant cell\\
 \hline
  $K_2$ & EC50 term for Treg inhibition of cytotoxic functions  \\
  \hline
  $K_3$ & EC50 term for how much TGF-$\beta$ each cell has \\
  \hline
  $K_4$ & EC50 term for TGF-$\beta$ activation of Tregs \\
  \hline
 $E_\text{NK}$ & rate of NKs clearing mutants  \\
  \hline
  $E_\text{CTL}$ & rate of CTLs clearing mutants \\
  \hline
  $\sigma_\text{NK}$ & NK source rate \\ 
  \hline
  $\sigma_\text{CTL}$ & CTL source rate per cleared mutant cell \\ 
  \hline
  $\sigma_\text{Treg}$ & Treg source rate per cleared mutant cell \\ 
  \hline
  $d_\text{NK}$ & NK death rate  \\ 
  \hline
  $d_\text{CTL}$ & CTL death rate \\ 
  \hline
  $d_\text{Treg}$ & Treg death rate \\ 
  \hline
  $k_\text{EMT}$ & EMT/MET rate  \\
  \hline
  $\sigma$ & standard deviation of noise in TGF-$\beta$ each cell receives  \\
  \hline
 $\tau_\text{max}$ & max amount of TGF-$\beta$ any cell can receive \\
  \hline 
 $\tau_\text{MUT}$ & rate of TGF-$\beta$ production by mutant cells\\
  \hline
 $\tau_\text{Treg}$ & rate of TGF-$\beta$ production by Treg\\
  \hline
\end{tabular}
  \caption{The model parameter names and descriptions. Note that many of these values are affected by the inflammation state of the system.}
\end{center}
\end{table}

@@

\begin{table}[H]
\begin{center}
 \begin{tabular}{|| c | p{7cm} | c | c ||} 
 \hline
 Name & Description & INFL Low Value & INFL High Value  \\ [0.5ex] 
 \hline\hline
  $p$ & weight of proliferation for tissue cells & 0.28 & \\ 
 \hline
 $d_C$  & weight of apoptosis for tissue cells & 0.14 & \\
 \hline
$\Delta_\text{MIE}$ & MIE & 0.6 & \\
 \hline
 $\Delta_\text{MGA}$ & MGA  & 0.2 &  \\
 \hline
  $\Delta_A$ & proportional decrease to weight of apoptosis for cells with mutated apoptosis pathway & 0.3 &  \\
  \hline
  $\Delta_\text{IE}$ & proportional increase to weight of immune evasion for cells with mutated immune evasion pathway & 0.48 &  \\
  \hline
  $\Delta_P$ & proportional increase to weight of proliferation for cells with mutated proliferation pathway & 0.36 &  \\
  \hline
 $K_0$ & EC50 term for negative feedback of tissue cells on own proliferation & 80 cells& \\
 \hline
 $K_1$ & EC50 term for probability of NK cell finding mutant cell & 8 cells& \\
 \hline
  $K_2$ & EC50 term for Treg inhibition of cytotoxic functions & 5 cells / volume & 0.025 cells / volume \\
  \hline
  $K_3$ & EC50 term for cumulative absorption of TGF-$\beta$ & 200 amount / volume & \\
  \hline
  $K_4$ & EC50 term for TGF-$\beta$ activation of Tregs & 50 amount / volume & \\
  \hline
 $E_\text{NK}$ & weight of NKs clearing mutants & 10 & 30  \\
  \hline
  $E_\text{CTL}$ & weight of CTLs clearing mutants & 200 & 600 \\
  \hline
  $\sigma_\text{NK}$ & NK source rate & 1.3 cells / cycle &  \\ 
  \hline
  $\sigma_\text{CTL}$ & CTL source rate per cleared mutant cell & 100 cells / (cleared mutants $\times$ cycles)& \\ 
  \hline
  $\sigma_\text{Treg}$ & Treg source rate per cleared mutant cell & 200 cells / (cleared mutants $\times$ concentration of TGF-$\beta$ $\times$ cycles)& \\ 
  \hline
  $d_\text{NK}$ & NK death rate & 0.13 / cycle &  \\ 
  \hline
  $d_\text{CTL}$ & CTL death rate & 0.0260 / cycle& \\ 
  \hline
  $d_\text{Treg}$ & Treg death rate & 0.0260 / cycle& \\ 
  \hline
  $k_\text{EMT}$ & EMT/MET rate & 0.01 / concentration of TGF-$\beta$ & \\
  \hline
  $\sigma$ & standard deviation of noise in TGF-$\beta$ each cell receives & 6 concentration of TGF-$\beta$ & \\
  \hline
 $\tau_\text{max}$ & max amount of TGF-$\beta$ any cell can receive & 500 concentration of TGF-$\beta$ &\\
  \hline 
 $\tau_\text{MUT}$ & rate of TGF-$\beta$ production by mutant cells & 0.05 concentration of TGF-$\beta$ / cell / cycle& \\
  \hline
 $\tau_\text{Treg}$ & rate of TGF-$\beta$ production by Treg & 0.5 concentration of TGF-$\beta$ / cell / cycle& \\
  \hline
  & RP Cancer Line  & 0.5 &  \\ 
 \hline
&  INFL High Duration & 30 cycles&   \\
 \hline
 & INFL Low Duration &30 cycles&   \\
 \hline
& Mes Threshold & 0.7 &    \\
 \hline
  & maximum initial mutation damage after warmup  & 0.01 &\\
  \hline
  & increase in probability to mutate for non-mutating proliferating cells & 0.0001 & \\
  \hline
\end{tabular}
  \caption{The model parameter names, descriptions, and values during both low and high inflammation. Parameters with only one value do not change with the inflammatory state.}
\end{center}
\end{table}

\newpage

\begin{table}[H]
\begin{center}
 \begin{tabular}{|| c | p{7cm} | c | c ||} 
 \hline
 Name & Description & INFL Low Value & INFL High Value  \\ [0.5ex] 
 \hline\hline
  $p$ & proliferation rate of tissue cells & 0.28 & \\ 
 \hline
 $d_C$  & death rate of tissue cells & 0.14 & \\
 \hline
$\Delta_\text{MIE}$ &  MIE & 0.6 & \\
 \hline
 $\Delta_\text{MGA}$ & MGA  & 0.2 &  \\
 \hline
  $\Delta_A$ & mutant cells decreased apoptosis & 0.3 &  \\
  \hline
  $\Delta_\text{IE}$ & mutant cells increased immune evasion & 0.48 &  \\
  \hline
  $\Delta_P$ & mutant cells increased proliferation & 0.36 &  \\
  \hline
 $K_0$ & EC50 term for negative feedback of tissue cells on own proliferation & 80 & \\
 \hline
 $K_1$ & EC50 term for probability of NK cell finding mutant cell & 8 & \\
 \hline
  $K_2$ & EC50 term for Treg inhibition of cytotoxic functions & 5 & 0.025  \\
  \hline
  $K_3$ & EC50 term for how much TGF-$\beta$ each cell has & 200 & \\
  \hline
  $K_4$ & EC50 term for TGF-$\beta$ activation of Tregs & 50 & \\
  \hline
 $E_\text{NK}$ & rate of NKs clearing mutants & 10 & 30  \\
  \hline
  $E_\text{CTL}$ & rate of CTLs clearing mutants & 200 & 600 \\
  \hline
  $\sigma_\text{NK}$ & NK source rate & 1.3 &  \\ 
  \hline
  $\sigma_\text{CTL}$ & CTL source rate per cleared mutant cell & 100 & \\ 
  \hline
  $\sigma_\text{Treg}$ & Treg source rate per cleared mutant cell & 200 & \\ 
  \hline
  $d_\text{NK}$ & NK death rate & 0.13 &  \\ 
  \hline
  $d_\text{CTL}$ & CTL death rate & 0.0260 & \\ 
  \hline
  $d_\text{Treg}$ & Treg death rate & 0.0260 & \\ 
  \hline
  $k_\text{EMT}$ & EMT/MET rate & 0.01 & \\
  \hline
  $\sigma$ & standard deviation of noise in TGF-$\beta$ each cell receives & 6 & \\
  \hline
 $\tau_\text{max}$ & max amount of TGF-$\beta$ any cell can receive & 500 &\\
  \hline 
 $\tau_\text{MUT}$ & rate of TGF-$\beta$ production by mutant cells & 0.05 & \\
  \hline
 $\tau_\text{Treg}$ & rate of TGF-$\beta$ production by Treg & 0.5 & \\
  \hline
  & RP Cancer Line  & 0.5 &  \\ 
 \hline
&  INFL High Duration & 30&   \\
 \hline
 & INFL Low Duration &30 &   \\
 \hline
& Mes Threshold & 0.7 &    \\
 \hline
  & maximum initial mutation damage after warmup  & 0.01 &\\
  \hline
  & increase in probability to mutate for non-mutating proliferating cells & 0.0001 & \\
  \hline
\end{tabular}
  \caption{The model parameter names, descriptions, and values during both low and high inflammation. Parameters with only one value do not change with the inflammatory state.}
\end{center}
\end{table}
